---
title: "Leverage Effects of Hedge Funds"
author: "jsn-jin @ UCLA"
date: "02-21-2020"
output:
  pdf_document:
    toc: yes
  html_document:
    highlight: tango
    number_sections: no
    theme: united
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Leverage Effects of Hedge Funds

Hedge funds can earn high profits by the use of leverage, but leverage also creates high risk. The simulations in this experiment explore the effects of leverage. Suppose a hedge fund owns \$1,000,000 of stock and used \$50,000 of its own capital and \$950,000 in borrowed money for the purchase. If the value of the stock falls below \$950,000 at the end of any trading day, then the
hedge fund must sell all the stock and repay the loan. This will wipe out its \$50,000 investment. **The hedge fund is said to be leveraged 20:1 since its position is 20 times the amount of its own capital invested.** Let $R_t$ denote the daily simple return of the stock at time t. Then $r_t = log(1 + R_t)$ is the daily log/cc return of this stock. Suppose $\{\cdots, r_{t-1}, r_t, r_{t+1}, \dots\}$ are independent normal random variables with mean 0.05/253 and variance $(0.23)^2 /253$, respectively.

**Suppose the hedge fund will sell the stock for a profit of at least \$100,000 if the value of the stock rises to at least \$1,100,000 at the end of one of the 45 trading days, sell it for a loss if the value falls below \$950,000 at the end of one of the 45 trading days, or sell after 45 trading days if the closing price has stayed between \$950,000 and \$1,100,000.**

## Part I

What is the probability that the value of the stock will be below
\$950,000 at the close of at least one of the next 45 trading days?

We can solve for the probability of interest by monte carlo simulation.

```{r}
mu.r <- 0.05 / 253
sigma.r <- 0.23 / sqrt(253)

n.day <- 100 # num of trading days
runs <- 5000
W0 <- 1000000
capital <- 50000

set.seed(2020)
r <- rnorm(runs * n.day, mu.r, sigma.r)
r <- matrix(r, nrow = n.day, byrow = TRUE)

dim(r)

head(r[, 1:5])
```

After generating the simulated data of cc returns, we can determine the simulated stock values.

```{r}
# compute the value of stock at the end of each day
# W_func <- function(r.col, w0 = W0, THRES = thres) {
#   res <- rep(-1, n.day) # a seq for storing results
#   res[1] <- w0 * exp(r.col[1])
#   hold <- n.day # by default, hold the stock till the last day
#   for (i in 2:n.day) {
#     if (res[i-1] >= THRES) {
#       res[i] <- res[i-1] * exp(r.col[i]) 
#     } else {
#       hold <- i-1
#       res[i:n.day] <- 0
#       break
#     }
#   }
#   list(res.col = res, hold.n = hold)
# }

W_func <- function(r.col, lb = W0 - capital, ub = W0 + 100000) {
  res <- rep(-1, n.day) # a seq for storing results
  res[1] <- W0 * exp(r.col[1])
  hold <- n.day # by default, hold the stock till the last day
  for (i in 2:n.day) {
    if (res[i-1] >= lb & res[i-1] <= ub) {
      res[i] <- res[i-1] * exp(r.col[i]) 
    } else if (res[i-1] > ub ) {
      hold <-  i - 1
      res[i:n.day] <- res[i-1]
      break
    } else {
      hold <- i - 1
      res[i:n.day] <- 0
      break
    }
  }
  list(res.col = res, hold.n = hold)
}

W <- matrix(-1, nrow = n.day, ncol = runs)
hold <- rep(-1, runs)

for (i in 1:runs) {
  temp <- W_func(r[,i])
  W[,i] <- temp$res.col
  hold[i] <- temp$hold.n
}

head(W[, 1:5])
```

Through monte carlo simulation, we can first calculate the probability that the stock value will be no less than \$950,000 at the close of each day. Then, subtracting this probability from 1 yields the probability of interest.
    
```{r}
ans1 <- 1 - sum(W[n.day,] >= W0 - capital) / runs
ans1
```

Moreover, we can consider the probability that the stock value will be below \$950,000 at the close of each day. 

```{r}
P <- rep(0, n.day)
for (i in 1:n.day) {
  P[i] <- sum(W[i,] <  W0 - capital)/runs
}

P
```

Though the probability $Pr(W_1 < 95,000)$ is nearly 1 as shown below. The probabilities of losing own capital grow quickly with trading days.

```{r}
x.vals <- seq(-0.25, 0.25, length = 1000)
df <- data.frame(x.vals, dnorm(x.vals, mu.r, sigma.r))
colnames(df) <- c("x", "y")
head(df$y)
library(ggplot2)
ggplot(data = df) + 
  geom_line(mapping = aes(y=y,x= x, color = "N( 0.05/253, (0.23)^2 /253)"), size = 1) + 
  labs(color = "Distribution", x = "x", y = "f(x)") +
  geom_vline(xintercept=log(0.95))
```


## Part II

What is the probability that the hedge fund will make a profit of at least \$100,000?

```{r}
sum(W[n.day,] > W0 + 100000) / runs
```


## Part III

What is the probability the hedge fund will suffer a loss?

```{r}
sum(W[n.day, ] < W0) / runs
```

The probability mirrors two possible events:
  
  - lose the stock at the end of any trading day
  - the stock value at the end of day 45 is between \$950,000 and \$1,000,000.

```{r}
ans2 <- sum(W[n.day, ] <= W0 & W[n.day, ] > W0 - capital) / runs
ans2 + ans1
```


## Part IV

What is the expected profit from this trading strategy?

```{r}
# rm(profit)
profit <- rep(0, runs)
for (i in 1:runs) {
  profit[i] <- W[hold[i], i] - W0
}
head(profit)
head(W[,1:6])
sprintf("The expected profit is %#.3f", mean(profit))
```


## Part V

What is the expected averaging return? When answering this question, remember that only \$50,000 was invested. Also, the units of return are time, e.g., one can express a return as a daily averaging return. Therefore, one must keep track of how long the hedge fund holds its position before selling.

The geometric average will be used to compute the average return.

```{r}
ret.avg <- rep(0, runs)
for (i in 1:runs) {
  ret.avg[i] <- ((profit[i]) / 50000) / hold[i]
}

head(ret.avg)
sprintf("The expected (arithmetic) average return is %#.4f", mean(ret.avg))

sprintf("Given the hedge fund makes a profit, the expected average return is %#.4f", 
        mean(ret.avg * (profit >= 0)) / mean(profit >= 0))

sprintf("The probability of making a profit is %#.4f", 
        mean(profit >= 0))

sprintf("Accordingly, the conditional profit is %f", 
        mean(profit * (profit >= 0)) / mean(profit >= 0))

sprintf("Given the hedge fund suffers a loss, the expected average return is %#.4f", 
        mean(ret.avg * (profit < 0)) / mean(profit < 0))

sprintf("The probability of making a profit is %#.4f",
        mean(profit < 0))

sprintf("Accordingly, the conditional profit is %f", 
        mean(profit * (profit < 0)) / mean(profit < 0))
```


