---
title: "Obtain Financial Data Online"
author: "jsn-jin @ UCLA"
date: "03-07-2020"
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Obtain Online Data

```{r}
library(BatchGetSymbols)

# Set dates
first.date <- as.Date("01/01/1998", format = "%m/%d/%Y")
last.date <- as.Date("12/31/2009", format = "%m/%d/%Y")
freq.data <- 'daily' # 'monthly'

# Set tickers
tickers <- c('^GSPC','SBUX') # The names of the stocks

l.out <- BatchGetSymbols(tickers = tickers, 
                         first.date = first.date,
                         last.date = last.date, 
                         freq.data = freq.data) # cache in tempdir()```

# This function is designed to make batch downloads of financial data using getSymbols. 
# Based on a set of tickers and a time period, the function will download the data for each ticker and return a report of the process, along with the actual data in the long dataframe format. 
# The main advantage of the function is that it automatically recognizes the source of the dataset from the ticker and structures the resulting data from different sources in the long format. A caching system is also presente, making it very fast.

summary(l.out) # See the elements in the list
str(l.out) # Structure of a list

SP <- subset(l.out$df.tickers, ticker == "^GSPC")
SBUX <- subset(l.out$df.tickers, ticker == "SBUX")

class(SP)
head(SP)
```

# Plot Historical Prices

```{r}
myplot <- function(df, name){
  library(dygraphs) # Interface to 'Dygraphs' Interactive Time Series Charting Library 
  library(xts) # eXtensible Time Series
  
  # Select a few columns
  data <- df %>% select(price.high, price.low, price.open, price.close) # Select data
  data <- as.xts(data, order.by = df$ref.date) # Transform the data into xts class
  
  dygraph(data, main = name) %>%
    dyCandlestick() %>%
    dyRoller(rollPeriod = 2) %>%
    dyRangeSelector()
}

myplot(SP, "S&P")
myplot(SBUX, "SBUX")
```



# Data Analysis

```{r}
myfun <- function(df, name) {
  df.n <- nrow(df)
  
  # CC return
  cc.ret <- diff(log(df$price.adjusted), lag = 1)
  
  # Time plot
  library(ggplot2)
  
  g <- ggplot() + 
    geom_line(aes(y = cc.ret, x = df$ref.date[2:df.n])) + 
    labs(x = "Time", y = "CC return") + 
    ggtitle(name)
  
  print(g)
  
  # Four panel plot
  par(mfrow = c(2,2))
  
	hist(cc.ret, main = "Monthly CC Returns",
	     xlab = name, probability = TRUE, col = "slateblue1")
	
	boxplot(cc.ret, outchar = TRUE, main = "Boxplot", col = "slateblue1")
	
	plot(density(cc.ret), type = "l", main = "Smoothed density",
	     xlab = "Monthly Return", ylab = "Density Estimate", col = "slateblue1")
	
	qqnorm(cc.ret, col = "slateblue1")
	qqline(cc.ret)
	
	# JB test
	library(quantmod)
	library(tseries)
	
	test <- jarque.bera.test(cc.ret)
	print(test)
	
	# Fit a GARCH(1, 1) to the time series
	x.arch <- garch(cc.ret, order = c(1,1), trace = FALSE)
  summary(x.arch)  # Diagnostic tests
  
  # Access estimates
  b <- x.arch$coef
  se <- sqrt(diag(x.arch$vcov))
  print(b)
  print(se)
}
```

We can apply the above function to get results of interest.

```{r}
myfun(SP, "S&P")
myfun(SBUX, "SBUX")
```

```{r}
# See [this webpage](https://seankross.com/2016/02/29/A-Q-Q-Plot-Dissection-Kit.html) for more on QQ plot.
```
